Dissecting the relationship between protein structure and sequence variation

### 2015-06-06 manuscript/voronoi/mnv.tex
MN update

    Throughout the previous sections, we carried out a comprehensive analysis and comparison of the most important structural determinants of sequence variability, using a dataset of $209$ monomeric enzymes. Examples of important sequence--structure relations include the correlations of measures of evolutionary rates such as {\it r4sJC} used in this work and sequence entropy, with measures of residue Contact Number (CN \& WCN) and Relative Solvent Accessibility (RSA). In addition, we have derived a new set of site-specific characteristics from the Voronoi Tessellation of protein 3D structures, some of which are capable of explaining sequence variability equally well or better than several structural quantities previously considered in the literature, such as B factor, RSA, and the traditional definitions of contact number and the weighted contact number (WCN) using $C_\alpha$ atomic coordinates (Figure \ref{fig:best_predictorER}).

    We also argued that the commonly used definitions of local packing density as measured by the Contact Number (CN) and the Weighted Contact Number (WCN), involve a free adjustable parameter in the definition. In order to get the best predictions of sequence variability, we fine-tuned the free parameters of the models for each proteins individually, and showed that the best performing parameter values of CN and WCN indicate a non-negligible influence of the long-range amino acid interactions, independently of the local packing density.

>>>>

    Throughout the previous sections, we carried out a comprehensive analysis and comparison of the most widely studied structural determinants of sequence variability, using a dataset of $209$ monomeric enzymes. Examples of important sequence--structure relations include the correlations of measures of evolutionary rates such as {\it r4sJC} used in this work and sequence entropy, with measures of residue Contact Number (CN \& WCN) and Relative Solvent Accessibility (RSA). In addition, we have derived a new set of site-specific characteristics from Voronoi Tessellation of protein 3D structure, some of which are capable of explaining sequence variability equally well or better than several structural quantities previously considered in the literature, such as B factor, RSA, and the traditional definitions of contact number and the weighted contact number (WCN) using $C_\alpha$ atomic coordinates (Figure \ref{fig:best_predictorER}).

    We also argued that the commonly used measures of local packing density, most importantly the Contact Number (CN) and the Weighted Contact Number (WCN), involve free adjustable parameters in the definitions which can be fine-tuned for each individual protein to obtain the best correlations between CN$/$WCN and sequence evolutionary rates. Although the best parameter values for CN \& WCN vary from one protein structure to another, the median values over the entire dataset indicate a non-negligible selection pressure due to long-range amino acid interactions on sequence evolution, independently of the local packing density.
